---
title: "Odds Generator"
author: "AwesomePie1453"
date: "4/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction

Back in Portfolio 7, I used an odds calculator to determine what the implied probabilities of future odds would be. In this lab, I want to adapt that, creating myself several calculators I can use for future use. I would like to be able to put in my own probabilities of an event to happen, and have it spit out the gambling odds for me. Here we go! 

```{r load}
library(implied)
library(tidyverse)
library(odds.converter)
```

### Start creating an odds calculator

As you can see on the original implied probability document I made, there are many different methods. I am just going to pick one and stick to it. I will be using the Margin Weights Proportion to the Odds method.

First, I have to add probabilies. I am going to use the first 5 games today in the MLB slate (4/22)

```{r Odds}
April_22_MLB_Odds <- rbind(c(2.50, 1.549),
                       c(1.80, 2.05),
                       c(2.56, 1.526),
                       c(2.05, 1.80),
                       c(1.741, 2.13))
colnames(April_22_MLB_Odds) <- c("Home", "Away")
rownames(April_22_MLB_Odds) <- c("Pirates vs Cubs", "Cardinals vs Reds", "Guardians vs Yankees", "Brewers vs Phillies", "Giants vs Nationals")
April_22_MLB_Odds
```
Now transform them into probabilities, using the Weights Proportion to the Odds method

```{r prob}
April22_MLB <- implied_probabilities(April_22_MLB_Odds, method = "wpo")
April22_MLB$probabilities
```

Now lets see if we can do the opposite.


```{r}
April_22_MLB_Prob <- rbind(c( 0.3772111, 0.6227889),
                       c(0.5338753,0.4661247),
                       c(0.3676585, 0.6323415))
colnames(April_22_MLB_Prob) <- c("Home", "Away")
rownames(April_22_MLB_Prob) <- c("Pirates vs Cubs", "Cardinals vs Reds", "Guardians vs Yankees")
April_22_MLB_Prob
```
I played with the margin until i found that .045 was the best with the wpo method.

```{r implied2}
April22_MLB_Implied<- implied_odds(
  April_22_MLB_Prob,
  method = "wpo",
  margin=0.045
  )
April22_MLB_Implied
```

Time to see if we can make a function to do this for any number we plug in. We will start with 40%. First lets put this, (as well as a value of 50%) into a matrix and get the betting odds

```{r Trial}
Trial <- rbind(c(0.40000, 0.60000),
                c(.5, .5))
colnames(Trial) <- paste0("Col", 1:2)
rownames(Trial) <- paste0("Row", 1:2)
Trial
```

```{r trial2}
Trial2 <- implied_odds(
  Trial,
  method = "wpo",
  margin=0.045
  )
Trial2
```

Perfect! Now, would we be able to isolate a particular value(s) from the matrix, and then make those american odds?

```{r trialMatrix}
Trial_Matrix <- Trial2$odds[c("Row1"),
                        c("Col1", "Col2")]
Trial_Matrix
```

```{r}
odds.dec2us(Trial_Matrix)
```


Let's run this backward to see if this worked!
```{r backward}
odds.us2dec(136.684)
odds.us2dec(-164.9007)
```


```{r probback}
Trial_Prob <- rbind(c(2.36684, 1.606426))
Trial_Prob2 <- implied_probabilities(Trial_Prob)
Trial_Prob2$probabilities
```
which is about 40% and 60%! slight differences, but oh well!
Noww.... to turn it into a function.

### Function

```{r function}
Prob_to_odds <- function(prob1, prob2){
  Probabilities <- rbind(c(prob1, prob2))
  Implied_Odds <- implied_odds(
  Probabilities,
  method = "wpo",
  margin=0.045
  )
  odds.dec2us(Implied_Odds$odds)
}
```

Which, if you run this with (.40, .60) you will get +137 and -165 rounded, which is what we got before! The function works!

I am also going to make a function for multiple games at once, and one for if the game has 3 outcomes

```{r multiple}
Prob_to_odds_multiple <- function(prob1, prob2, prob3, prob4, prob5, prob6){
  Probabilities <- rbind(c(prob1, prob2),
                         c(prob3, prob4),
                         c(prob5, prob6))
  Implied_Odds <- implied_odds(
  Probabilities,
  method = "wpo",
  margin=0.045
  )
  odds.dec2us(Implied_Odds$odds)
}
```

```{r three}
Prob_to_odds_three <- function(prob1, prob2, prob3){
  Probabilities <- rbind(c(prob1, prob2, prob3))
  Implied_Odds <- implied_odds(
  Probabilities,
  method = "wpo",
  margin=0.045
  )
  odds.dec2us(Implied_Odds$odds)
}
```

One last thing Ill do despite spending way too much time on this, I'm going to create a function that allows me to see the difference between my projected odds, and the actual bookmaker odds. 

```{r comparison}
Odds_comparison <- function(MyProb1, MyProb2, BookOdds1, BookOdds2){
  Probabilities <- rbind(c(MyProb1, MyProb2))
  colnames(Probabilities) <- paste0("Col", 1:2)
  rownames(Probabilities) <- paste0("Row", 1)
  Implied_Odds <- implied_odds(
  Probabilities,
  method = "wpo",
  margin=0.045
  )
  MyOdds <- odds.dec2us(Implied_Odds$odds)
  MyOdds2 <- MyOdds[c("Row1"),
                    c("Col1")]
  MyOdds3 <- MyOdds[c("Row1"),
                    c("Col2")]
  Difference1 <- BookOdds1 - MyOdds2
  Difference2 <- BookOdds2 - MyOdds3
  print(Difference1)
  print(Difference2)
}
```

Now that this all works. Let's try this last one out. Tomorrow, at 4:10pm, The Toronto Blue Jays and starting pitcher Alek Manoah take on Jose Urquidy and the Houston AStros. I believe that there is a 53% chance that the Blue Jays win tomorrow. The betting odds are -125 for the Jays. Lets plug in these numbers and see how far I was off

```{r JaysAstros}
Odds_comparison(.53, .47, -125, 105)
```
Wow, only 1 to 2 cents? thats pretty crazy. I'm going to double check the numbers of my percentages with my original function

```{r JaysTest}
Prob_to_odds(.53, .47)
```

-123 and +103, as opposed to -125 and +105. Crazy. Go me.


Well that's it for this portfolio! Took me a long time but really happy with it. Only thing i could change is I will probably make the funciton names easier to type. But thats a problem for another day. Thanks!