---
title: "Odds Generator"
author: "AwesomePie1453"
date: "4/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction

Back in Portfolio 7, I used an odds calculator to determine what the implied probabilities of future odds would be. In this lab, I want to adapt that, creating myself several calculators I can use for future use. I would like to be able to put in my own probabilities of an event to happen, and have it spit out the gambling odds for me. Here we go! 

```{r load}
library(implied)
library(tidyverse)
library(odds.converter)
```

### Start creating an odds calculator

As you can see on the original implied probability document I made, there are many different methods. I am just going to pick one and stick to it. I will be using the Odds Ratio Method.

First, I have to add probabilies. I am going to use the first 5 games today in the MLB slate (4/22)

```{r Odds}
April_22_MLB_Odds <- rbind(c(2.50, 1.549),
                       c(1.80, 2.05),
                       c(2.56, 1.526),
                       c(2.05, 1.80),
                       c(1.741, 2.13))
colnames(April_22_MLB_Odds) <- c("Home", "Away")
rownames(April_22_MLB_Odds) <- c("Pirates vs Cubs", "Cardinals vs Reds", "Guardians vs Yankees", "Brewers vs Phillies", "Giants vs Nationals")
April_22_MLB_Odds
```
Now transform them into probabilities, using the Odds Ratio Method

```{r prob}
April22_MLB <- implied_probabilities(April_22_MLB_Odds, method = "or")
April22_MLB$probabilities
```

